% "World of Adventure" campaign guide - Skills
% Written by Christopher Thomas.

\chapter{Skills}
\label{sect-skills}

\section{Modified Skills}

%
\subsection{Diplomacy}

The Diplomacy skill may be used on groups of people, not just individuals.
Modifications to the DC are at the DM's discretion.
%

%
\subsection{Knowledge (arcana)}

This is the study of what kinds of things \textit{can} happen, cataloguing
the kinds of creatures and magical oddities that have been observed in the
field. This also covers the more philosophical end of \textit{why} they
happen (with Spellcraft covering the more practical end).

At the DM's discretion, having lots of ranks in Knowledge (arcana) may
provide a synergy bonus to appropriate Spellcraft rolls.
%

%
\subsection{Knowledge (dungeoneering)}

This is re-skinned as being the catch-all course given to adventurers and
similar individuals who need some knowledge but don't have the time or
skill points to get in-depth training.

For rolls where the DM feels that it is appropriate, Knowledge
(dungeoneering) can substitute for skills such Knowledge (arcana),
Knowledge (nature), Knowledge (geography), or Survival. The DC is
typically 5 points higher than it would be if you actually had the
relevant skill.
%

%
\subsection{Knowledge (engineering)}

This is split into two skills:

\begin{itemize}
%
\item \textbf{Knowledge (civil engineering)} covers buildings, as well as
structures such as bridges, paved roads, sewer systems, and so forth.
%
\item \textbf{Knowledge (mechanical engineering} covers machinery, from
siege weapons to clockwork mechanisms.
%
\end{itemize}

These will normally be combined with appropriate Craft or Profession
skills (such as Craft (mechanisms), Craft (siege weapons), Profession
(architect), etc).
%

%
\subsection{Use Magic Device}

If you're an arcane caster trying to use a scroll from a different arcane
caster class (such as a bard trying to use a wizard scroll), the DC is
10~+~spell~level.

If you're an arcane or divine caster and are trying to use a scroll from
the other type (arcane or divine), the DC is 15~+~spell~level.

If you're a non-caster (such as a rogue) and are trying to use a scroll,
the DC is 20~+~spell~level.

``Spell completion'' devices (like wands) have the same requirements.
%

%
%
\section{New Skills}

%
\subsection{Craft (mechanisms)}

This is a widely-known skill used for producing machinery (anything from
factory looms driven by water power to treadle-operated sewing machines).

If constructs are part of the setting, this skill is also used for
creating a construct's body.
%

%
% This is the end of the file.
