% "World of Adventure" campaign guide - Feats
% Written by Christopher Thomas.

\chapter{Feats}
\label{sect-feats}

\section{Modified Feats}

%
\subsection{Divine Interference}

Prerequisite: Cleric 5

You can sacrifice a spell slot (of 1st level or higher) to force a re-roll
of an attack roll or skill check, friend or foe, made within 30 feet. There
is a bonus or penalty to the roll (your choice) equal to the level of the
spell you sacrificed.
%

%
\subsection{Item Creation Feats}

The magic item creation feats' level requirements are reduced, since most
NPCs die of old age before reaching level 7:

\begin{tabular}{l|l}\hline
\textbf{Level 1} & Scribe Scroll, Brew Potion \\
\textbf{Level 3} & Craft Wondrous Item, Craft Magical Arms and Armour \\
\textbf{Level 5} & Craft Wand, Forge Ring, Craft Lesser Golem \\
\textbf{Level 7} & Craft Rod, Craft Staff, Craft Greater Golem \\
\end{tabular}

The most immediate consequence is that a starting wizard can take Brew
Potion instead of Scribe Scroll as a starting feat, if desired.

See the chapter on magic item creation for details about when each type
of object would be suitable.

For golem feats, see the section on construct house rules. Constructs are
an optional part of the campaign setting; the DM may prohibit them, or may
adjust the price to make them rare (per the costs table in that section).
%

%
\subsection{Leadership}

Prerequisite: character level 3

You are the type of person that others want to follow or want to work for.
You do not automatically get staff or followers, but you have an easier
time hiring them and managing them.

See the section on employees for a detailed discussion of working in groups.

If you have the Leadership feat, you can manage three additional employees
while trying to do your own work at the same time, or double the number
of employees (including doubling your wisdom bonus) if managing full-time.

You also get a bonus of +3 to the rolls to find new employees and keeping
your existing employees.
%

%
\subsection{Power Attack}

Secondary attacks only get half the damage bonus from Power Attack, rather
than the full bonus.

This mostly affects monsters that have the Power Attack feat.
%

%
%
\section{New Feats}

%
\subsection{Arcane Training}

Prerequisite: Int 10+

You have basic training in arcane spellcasting without being a career
wizard. You know three wizard cantrips (and only those three), which you
can prepare and cast as a wizard would.

Taking this feat multiple times gives more cantrips known but does not
provide more spell slots. Multiclassing into wizard likewise only provides
the total number of cantrip spell slots you'd get as a wizard.

\textit{Adapted from Tony's Duanor campaign, with permission.}
%

%
\subsection{Combat Awareness}

Prerequisite: Combat Expertise, Wis 13+

You are able to quickly recognize dangerous situations. If you are
surprised in combat, you are only considered flat-footed by opponents who
beat your initiative roll.

\textit{Adapted from Tony's Duanor campaign, with permission.}
%

%
\subsection{Craft Golem (Greater or Lesser)}

Prerequisite: Wizard, Int 13+ (Lesser) or 15+ (Greater)

The Craft Lesser Golem feat represents specialized training in the
manufacture of constructs. This is what lets you design constructs and
cast the animation enchantment.

The Craft Greater Golem feat lets you make constructs more easily, gives
you more build points for a given size, and lets you build constructs with
animal intelligence.

See the section on construct house rules for further information.
%

%
\subsection{Formation Fighting}

Prerequisite: base attack bonus +1

When leading a formation, you receive a +2 bonus to all morale rolls
involving your troops. In addition, if a simple majority of the unit's
members have this feat, the following effects apply:
\begin{itemize}
\item All unit members receive a +1 bonus to attack rolls.
\item The front rank receives a +1 bonus to AC.
\item Ranks behind the front rank receive a +4 bonus to AC.
\end{itemize}

\textit{Adapted from Tony's Duanor campaign, with permission.}

\textit{NOTE - This was intended for use with Tony's mass combat system.
It still works well enough with my system (one roll for each group of
combatants).}
%

%
\subsection{Inspired Artist}

Prerequisites: appropriate Craft or Profession 1, Wis 13+

You are gifted in creative works. When you make a Craft or Profession roll
involving creation of a new composition (unique and with artistic merit),
you receive a +2 bonus to the Craft or Profession roll, and you have a
higher-than-normal income multiplier (exact value at the DM's discretion;
typically doubling income for an early-career character).

\textit{Adapted from Tony's Duanor campaign, with permission.}
%

%
\subsection{Martial Tutelage}

Prerequisite: base attack bonus +2, Formation Fighting,
suitable noble title or military rank

You are expected to be able to wage war and lead troops in a military
campaign, and you have been extensively trained in doing so. You receive
a +2 competency bonus to checks regarding leading troops (such as morale
checks), stacking with the bonus from Formation Fighting (if individual
unit leaders have it). In addition, as long as you are actively directing
your forces (i.e. lines of communication are open), all forces under your
control receive a +2 bonus to their initiative.

\textit{Adapted from Tony's Duanor campaign, with permission.}

\textit{NOTE - This was intended for use with Tony's mass combat system.
It still works well enough with my system (one roll for each group of
combatants).}
%

%
\subsection{Nature Training}

Prerequisite: Wis 10+

You have basic training in druidic spellcasting without being a career
druid. You know three druid orisons (and only those three), which you can
prepare and cast as a druid would.

Taking this feat multiple times gives more orisons known but does not
provide more spell slots. Multiclassing into druid likewise only provides
the total number of orison spell slots you'd get as a druid.

\textit{Adapted from Tony's Duanor campaign, with permission.}
%

%
\subsection{Performer Training}

Prerequisites: Perform 1, Cha 13+

Your extensive training in the performing arts, history, and culture
gives you a +2 bonus on Performance and Linguistics checks. You also gain
1 additional language point.

\textit{Adapted from Tony's Duanor campaign, with permission.}
%

%
\subsection{Puglist}

Prerequisites: base attack bonus +1

You have trained in boxing or other unarmed combat, and can absorb blows
that most others wouldn't. You have an addtional 2 points of damage
reduction against nonlethal bludgeoning damage (such as from unarmed combat
or most nonlethal weapons).

\textit{Adapted from Tony's Duanor campaign, with permission.}
%

%
\subsection{Sharpshooter}

Prerequisites: Dex 13+, Precise Shot

You have uncannily good aim with missile weapons. Targets strick by your
arrows or quarels take additional damage equal to your dexterity modifier.

\textit{Adapted from Tony's Duanor campaign, with permission.}
%

%
\subsection{Skill Specialization}

Prerequisites: Knowledge or Craft skill 1, Skill Focus on that skill

You have specialized knowledge within a field of study (such as heraldry
for Knowledge (nobility)), or specialized expertise within a craft (such as
daggers for Craft (weaponsmith)). When you make a skill check within this
field of expertise, you get a +2 bonus with an additional +1 for every
5 full skill ranks (so, +3 at 5 ranks, +4 at 10 ranks, etc). This stacks
with Skill Focus.

\textit{Adapted from Tony's Duanor campaign, with permission.}
%

%
\subsection{Superior Unarmed Strike}

Prerequisites: Base attack +4, Improved Unarmed Strike

Your unarmed strikes are particularly devastating. The base damage for
your unarmed strikes is 1d6 (rather than 1d3).

\textit{Adapted from Tony's Duanor campaign, with permission.}
%

%
\subsection{Theurgist Training}

Prerequisite: Wis 10+

You have basic training in divine spellcasting without being a career
cleric. You know three cleric orisons (and only those three), which you
can prepare and cast as a cleric would.

Taking this feat multiple times gives more orisons known but does not
provide more spell slots. Multiclassing into cleric likewise only provides
the total number of orison spell slots you'd get as a cleric.

\textit{Adapted from Tony's Duanor campaign, with permission.}
%

%
\subsection{Tricky Combatant}

Prerequisite: Str 13+ \textit{or} Dex 13+

You are particularly good at applying strength, speed, or both to combat
maneuvers (grappling, disarming, etc). You receive a +2 bonus to your
Combat Maneuver Bonus and Combat Maneuver Defense for combat maneuvers of
all types.

\textit{Adapted from Tony's Duanor campaign, with permission.}
%

%
\subsection{Wilderness Training}

Prerequisite: Kn: Nature 1

You have received extensive practical training on surviving in the
wilderness. As a result, you receive a +2 bonus on Knowledge (nature) and
Survival checks.

\textit{Adapted from Tony's Duanor campaign, with permission.}
%


%
% This is the end of the file.
